Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-05-21 doc/thesis/methods_results.tex
Most of Alex's changes done

<<setup, include=FALSE>>=
    source("global.R")
@

@@

To test the full kernel-assisted \gls{ABC} algorithm, we simulated three trees each under
a variety of parameter values, and ran the \software{netabc} program to
estimate posterior distributions for the parameters. The parameter values and
priors used are listed in \cref{tab:abcexpt}. The tree kernel meta-parameters
were set to $\lambda = 0.3$ and $\sigma = 4$. The \gls{SMC} algorithm was run
with 1000 particles, five sampled datasets per particle, and the $\alpha$
parameter (not to be confused with the \gls{BA} preferential attachment
parameter, see \cref{subsec:adaptsmc}) set to 0.95. The algorithm was
stopped when the acceptance rate of the \gls{MH} kernel dropped below 1.5\%,
the same criterion used by \citeauthor{del2012adaptive}. Approximate marginal
posterior densities for each parameter were calculated using the
\software{density} function in \software{R} applied to the final weighted
population of particles. {\color{red}\sout{Credible intervals were obtained for
each parameter using the \software{HPDinterval} function in the \software{coda}
package~\autocite{plummer2006coda}.}} {\color{blue}\uline{Posterior means
obtained for each parameter using the \software{wtd.mean} function in the
\software{Hmisc} package~\autocite{harrell2016hmisc}. Credible intervals were
obtained using the \software{hpd} function in the \software{TeachingDemos}
package~\autocite{snow2013teachingdemos}.}}

@@

<<treestats, include=FALSE>>=
    params <- c("alpha", "I", "m", "N")

    get.treestats <- function (param) {
        stats <- paste0("../../simulations/kernel-", param, "/statistics/*")
        d <- collect.data(stats)
        trees <- paste0("../../simulations/kernel-", param, "/tree/*")
        m <- collect.metadata(trees)
        rownames(m) <- sub("../../simulations/", "", rownames(m))
        d <- merge(d, m, by=0, suffixes=c("", ".1"))
        if (param == "alpha") {
            d <- subset(d, m == 2)
            colnames(d)[colnames(d) == "nsimnode"] <- "I"
        }
        setDT(d)
    }

    alpha.stats <- get.treestats("alpha")
    I.stats <- get.treestats("I")
    m.stats <- get.treestats("m")
    N.stats <- get.treestats("N")

    alpha.sackin.cor <- alpha.stats[,cor.test(alpha, sackin, method="spearman")]
    I.sackin.cor <- I.stats[,cor.test(I, sackin, method="spearman")]
    m.sackin.cor <- m.stats[,cor.test(m, sackin, method="spearman")]
    N.sackin.cor <- N.stats[,cor.test(N, sackin, method="spearman")]
    stopifnot(alpha.sackin.cor$p.value < 0.05)
    stopifnot(I.sackin.cor$p.value < 0.05)
    stopifnot(m.sackin.cor$p.value < 0.05)
    stopifnot(N.sackin.cor$p.value < 0.05)

    alpha.ratio.cor <- alpha.stats[,cor.test(int.tip.ratio, alpha)]
    I.ratio.cor <- I.stats[,cor.test(int.tip.ratio, I)]
    m.ratio.cor <- m.stats[,cor.test(int.tip.ratio, m)]
    N.ratio.cor <- N.stats[,cor.test(int.tip.ratio, N)]
    stopifnot(alpha.ratio.cor$p.value < 1e-5)
    stopifnot(I.ratio.cor$p.value < 1e-5)
    stopifnot(m.ratio.cor$p.value < 1e-5)
    stopifnot(N.ratio.cor$p.value < 1e-5)
@

Trees simulated under different values of \gls{alpha} were visibly quite
distinct (\cref{fig:alphatrees}). In particular, higher values of \gls{alpha}
produce networks with a small number of highly connected nodes which, once
infected, are likely to transmit to many other nodes. This results in a more
unbalanced, ladder-like structure in the phylogeny, compared to networks with
lower \gls{alpha} values. None of the other three parameters produced trees
which were as easily distinguished from each other
(\cref{fig:Itrees,fig:mtrees,fig:Ntrees,fig:Itrees}).  Sackin's index, which
measures tree imbalance, was significantly correlated with all four parameters
    (for $\alpha$, $I$, $m$, and $N$ respectively: Spearman's rho =
     \Sexpr{round(alpha.sackin.cor$estimate, 2)},
     \Sexpr{round(I.sackin.cor$estimate, 2)},
     \Sexpr{round(m.sackin.cor$estimate, 2)},
     \Sexpr{round(N.sackin.cor$estimate, 2)};
     $p$-values
     $\Sexpr{pp(alpha.sackin.cor$p.value)}$,
     $\Sexpr{pp(I.sackin.cor$p.value)}$,
     $\Sexpr{pp(m.sackin.cor$p.value)}$,
     $\Sexpr{pp(N.sackin.cor$p.value)}$).
The ratio of internal to terminal branch lengths was negatively correlated with
\gls{alpha} and \gls{I}, and positively correlated with \gls{m} and \gls{N}
  (Spearman's rho
    \Sexpr{round(alpha.ratio.cor$estimate, 2)},
    \Sexpr{round(I.ratio.cor$estimate, 2)},
    \Sexpr{round(m.ratio.cor$estimate, 2)},
    \Sexpr{round(N.ratio.cor$estimate, 2)};
  all $p < 10^{-5}$).

@@

<<classifiers, include=FALSE>>=
    get.kernel.data <- function (param, step) {
        md <- collect.metadata(paste0('../../simulations/kernel-', param, '/', step, '/*'))
        if ('rbf_variance' %in% colnames(md)) {
          md <- subset(md, rbf_variance == 4 & decay_factor == 0.3)
        }
        if (param == 'alpha') {
            md <- subset(md, m == 2)
        }
        d <- setDT(collect.data(rownames(md)))
        if (param %in% c('alpha', 'm')) {
            d[,list(rsquared=mean(rsquared)), by=c('nsimnode', 'ntip')]
        } else if (param == 'I') {
            d[,list(rsquared=mean(rsquared)), by=ntip]
        } else {
            d[,list(rsquared=mean(rsquared)), by=c('I', 'ntip')]
        }
    }
    kernel.alpha <- get.kernel.data('alpha' ,'classifier')
    kernel.m <- get.kernel.data('m', 'classifier')
    kernel.I <- get.kernel.data('I', 'classifier')
    kernel.N <- get.kernel.data('N', 'classifier')

    sackin.alpha <- get.kernel.data('alpha', 'stats-classifier')
    sackin.m <- get.kernel.data('m', 'stats-classifier')
    sackin.I <- get.kernel.data('I', 'stats-classifier')
    sackin.N <- get.kernel.data('N', 'stats-classifier')

    nltt.alpha <- get.kernel.data('alpha', 'nltt-classifier')
    nltt.m <- get.kernel.data('m', 'nltt-classifier')
    nltt.I <- get.kernel.data('I', 'nltt-classifier')
    nltt.N <- get.kernel.data('N', 'nltt-classifier')
@

@@

When classifying $\alpha$, the \gls{kSVR} classifier had an average $R^2$ of 
    \Sexpr{kernel.alpha[,round(mean(rsquared), 2)]},
compared to 
    \Sexpr{nltt.alpha[,round(mean(rsquared), 2)]}
for the \gls{nltt}-based SVR, and
    \Sexpr{sackin.alpha[,round(mean(rsquared), 2)]}
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify the $m$ parameter in any epidemic scenario, with average
$R^2$ values of 
  \Sexpr{kernel.m[,round(mean(rsquared), 2)]} for \gls{kSVR},
  \Sexpr{nltt.m[,round(mean(rsquared), 2)]} for the \gls{nltt}, and
  \Sexpr{sackin.m[,round(mean(rsquared), 2)]}
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVR} was slightly higher
on 1000-tip trees 
    (average $R^2$ 
     \Sexpr{kernel.m[ntip == 100,round(mean(rsquared), 2)]},
     \Sexpr{kernel.m[ntip == 500,round(mean(rsquared), 2)]},
     \Sexpr{kernel.m[ntip == 1000,round(mean(rsquared), 2)]}
     for 100, 500, and 1000 tips respectively).

The accuracy of classifiers $I$ varied significantly with the number of tips in
the tree. For 100-tip trees, the average $R^2$ values were
  \Sexpr{kernel.I[ntip == 100, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 100, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 100, round(mean(rsquared), 2)]}
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  \Sexpr{kernel.I[ntip == 500, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 500, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 500, round(mean(rsquared), 2)]}.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  \Sexpr{kernel.N[,round(min(rsquared), 2)]}
for the smallest epidemic and smallest sample size, to
  \Sexpr{kernel.N[,round(max(rsquared), 2)]}
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  \Sexpr{nltt.N[,round(min(rsquared), 2)]}
to
  \Sexpr{nltt.N[,round(max(rsquared), 2)]}.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  \Sexpr{sackin.N[,round(mean(rsquared), 2)]}
and little variation between scenarios.

@@

<<gridsearch, include=FALSE>>=
    d <- setDT(collect.data("../../simulations/aggregates/gridsearch/*"))
    d[,error := abs(true_value - point.est)]
    setkey(d, parameter)

    # correlation between number of tips and error
    tip.cor <- d[,cor.test(error, tips, method="spearman")[c("p.value", "estimate")], by=parameter]
    setkey(tip.cor, parameter)
    stopifnot(tip.cor["alpha", p.value] < 0.05)
    stopifnot(tip.cor["I", p.value] < 0.05)
    stopifnot(tip.cor["N", p.value] < 0.05)
    stopifnot(tip.cor["m", p.value] > 0.05)

    # CI takes up >75% of grid
    stopifnot(d["alpha", all(upper - lower >= 0.75*2)])
    stopifnot(d["I", all(upper - lower >= 0.75*5000)])
    stopifnot(d["N", all(upper - lower >= 0.75*15000)])
    stopifnot(d["m", all(upper - lower >= 0.75*6)])

    # correlation between value and error
    av <- d[,list(p=summary(aov(error~factor(true_value)))[[1]][["Pr(>F)"]][1]),
            by=parameter]
    stopifnot(av["alpha", p] < 0.05)
    stopifnot(av["I", p] < 0.05)
    stopifnot(av["m", p] < 0.05)
    stopifnot(av["N", p] < 0.05)

    alpha.test <- d["alpha", wilcox.test(.SD[true_value %in% c(1.0, 1.25),error], 
                                         .SD[!true_value %in% c(1.0, 1.25),error])]
    stopifnot(alpha.test$p.value < 0.05)

    I.test <- d["I", wilcox.test(.SD[true_value >= 2000 & true_value <= 3000, error], 
                                 .SD[true_value < 2000 | true_value > 3000, error])]
    stopifnot(I.test$p.value < 0.05)

    m.test <- d["m", wilcox.test(.SD[true_value == 1, error], 
                                 .SD[true_value > 1, error])]
    stopifnot(m.test$p.value < 0.05)

    N.test <- d["N", wilcox.test(.SD[true_value <= 3000, error], 
                                 .SD[true_value > 3000, error])]
    stopifnot(N.test$p.value < 0.05)

    value.cor <- d[,cor.test(error, true_value)[c("p.value", "estimate")], by=parameter]
    setkey(value.cor, parameter)
    stopifnot(value.cor["I", p.value] >= 0.05)
@

The accuracy of grid search estimates largely paralleled that of the \gls{kSVR}
classifiers. \Cref{fig:gridest} shows point estimates and 95\% highest density
intervals for each of the \gls{BA} parameters, for one replicate experiment
with 500-tip trees. Plots showing the point estimates for all replicates can be
found in \cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all
parameters except $m$, the error of point estimates was negatively correlated
with the number of sampled tips in the tree (for
\gls{alpha}, \gls{I}, and \gls{N} respectively: Spearman's $\rho$ = 
    \Sexpr{tip.cor["alpha", round(estimate, 2)]},
    \Sexpr{tip.cor["I", round(estimate, 2)]},
    \Sexpr{tip.cor["N", round(estimate, 2)]};
$p$-values
    $\Sexpr{tip.cor["alpha", pp(p.value)]}$,
    $\Sexpr{tip.cor["I", pp(p.value)]}$,
    $\Sexpr{tip.cor["N", pp(p.value)]}$).
The highest density intervals obtained for all parameters were extremely wide,
occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

The \gls{alpha} parameter was the most accurately estimated, with point
estimates having an average deviation of 
    \Sexpr{d["alpha", round(mean(error), 2)]}
from the true value, on a grid from 0 to 2. The error of point estimates varied
significantly between true values of \gls{alpha}
    (one-way \gls{ANOVA}, $p$ $\Sexpr{pp(av["alpha", p], eq=TRUE)}$). In
particular, errors were lower for the values \gls{alpha} = 1.0 and 1.25 than
for the other values
    (average errors 
    \Sexpr{d["alpha", .SD[true_value %in% c(1.0, 1.25), round(mean(error), 2)]]}
    for \gls{alpha} = 1.0 or 1.5 vs.
    \Sexpr{d["alpha", .SD[!true_value %in% c(1.0, 1.25), round(mean(error), 2)]]}
    for \gls{alpha} $\neq$ 1.0 or 1.5),
and this difference was significant
    (Wilcoxon rank-sum test, $p \Sexpr{pp(alpha.test$p.value, eq=TRUE)}$,
     \cref{fig:gridptalpha}).
These two values exhibited different qualitative behaviour than the other
values in terms of the distribution of kernel scores along the grid
(\cref{fig:gridalpha}). In particular, there was a pronounced peak in scores
around the true value, in contrast to the other values where the scores were
flat around the true value. The effect was most obvious for the value
\gls{alpha} = 1.25.

The average absolute error of the point estimates for \gls{I} was 
    \Sexpr{d["I", round(mean(error))]} individuals,
on a grid of 500 to 5000, and these errors differed between true values of
\gls{I}
    (one-way \gls{ANOVA}, $p \Sexpr{pp(av["I", p], eq=TRUE)}$).
The errors for $2000 \leq I \leq 3000$ were higher than those for the other
values
    (average errors
     \Sexpr{d["I", .SD[true_value >= 2000 & true_value <= 3000, round(mean(error))]]}
     for $2000 \leq I \leq 3000$ vs.
     \Sexpr{d["I", .SD[true_value < 2000 | true_value > 3000, round(mean(error))]]}
     for $I < 2000$ or $I > 3000$),
and this difference was significant
    (Wilcoxon rank-sum test, $p \Sexpr{pp(I.test$p.value, eq=TRUE)}$,
     \cref{fig:gridptI}).
Kernel score distributions for all test values exhibited a similar rounded
shape (\cref{fig:gridI}).

The average error for \gls{m} was
    \Sexpr{d["m", round(mean(error), 2)]} edges per vertex,
on a grid from 1 to 6. The error varied significantly between the true values
of \gls{m} 
    (one-way \gls{ANOVA}, $p \Sexpr{pp(av["m", p], eq=TRUE)}$).
Errors for the value \gls{m} = 1 were lower than the other values
    (average errors
    \Sexpr{d["m", .SD[true_value == 1, round(mean(error), 2)]]}
    for $m = 1$ vs.
    \Sexpr{d["m", .SD[true_value > 1, round(mean(error), 2)]]}
    for $m > 1$),
and this difference was significant
    (Wilcoxon rank-sum test, $p \Sexpr{pp(m.test$p.value, eq=TRUE)}$,
     \cref{fig:gridptm}).
The value $m = 1$ causes the network to take on a distinct shape relative to
higher \gls{m} values, namely a tree (\ie there are no cycles, see
\cref{subsec:treeshape}). The kernel score distribution had a peak at $m = 1$
when this was the true value, and a valley at $m = 1$ when the true value of
$m$ was greater that 1 (\cref{fig:gridm}).

The average error for \gls{N} was 
    \Sexpr{d[parameter == "N", round(mean(error))]} individuals,
on a grid from 1000 to 15000, and was varied significantly with the true value
of \gls{N}
    (one-way \gls{ANOVA}, $p \Sexpr{pp(av["N", p], eq=TRUE)}$).
The errors were lower for $N \leq 3000$
    (average errors
     \Sexpr{d["N", .SD[true_value <= 3000, round(mean(error))]]}
     for $N \leq 3000$ vs.
     \Sexpr{d["N", .SD[true_value > 3000, round(mean(error))]]}
     for $N > 3000$),
and this difference was significant
    (Wilcoxon rank-sum test, $p \Sexpr{pp(N.test$p.value, eq=TRUE)}$,
     \cref{fig:gridptN}).
The kernel score distribution had a peak at $N = 1000$ when this was the true
value, and a valley there otherwise (\cref{fig:gridN}). Except for this valley,
the distributions were flat for $N > 3000$.

@@

<<point_est, include=FALSE>>=
    f <- Sys.glob("../../simulations/abc-pa-free-m/point-estimate/*")
    d <- fread(f)
    d[,m := floor(m)]
    d[,alpha_error := abs(true_alpha - alpha)]
    d[,N_error := abs(true_N - N)]
    d[,I_error := abs(true_I - I)]
    d[,m_error := abs(true_m - m)]

    alpha.av <- anova(lm(alpha_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))
    m.av <- anova(lm(m_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))
    N.av <- anova(lm(N_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))
    I.av <- anova(lm(I_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))

    m.tbl <- d[,prop.table(table(m_error))]
    zero.alpha.test <- wilcox.test(d[true_alpha == 0, alpha_error], 
                                   d[true_alpha != 0, alpha_error])
    alpha.I.test <- wilcox.test(d[true_alpha < 1, I_error],
                                d[true_alpha >= 1, I_error])
    I.I.test <- wilcox.test(d[true_I == 1000, I_error],
                            d[true_I == 2000, I_error])
    alpha.m.test <- wilcox.test(d[true_alpha == 0 | true_alpha == 1, m_error],
                                d[true_alpha == 0.5 | true_alpha == 1.5, m_error])
    #stopifnot(alpha.av["factor(true_m)", "Pr(>F)"] <= 0.05)
    #stopifnot(alpha.av["factor(true_I)", "Pr(>F)"] > 0.05)
    #stopifnot(I.I.test$p.value < 1e-5)
    #stopifnot(I.av["factor(true_m)", "Pr(>F)"] > 0.05)
    #stopifnot(m.av["factor(true_m)", "Pr(>F)"] > 0.05)
    #stopifnot(m.av["factor(true_I)", "Pr(>F)"] > 0.05)
    #stopifnot(m.av["factor(true_alpha)", "Pr(>F)"] > 0.05)
    #stopifnot(min(N.av[1:3,"Pr(>F)"]) > 0.05)

    # one-way ANOVAs
    stopifnot(anova(lm(alpha_error~true_m, d))[1, "Pr(>F)"] > 0.05)
    stopifnot(anova(lm(alpha_error~factor(true_I), d))[1, "Pr(>F)"] > 0.05)
    stopifnot(anova(lm(I_error~true_m, d))[1, "Pr(>F)"] > 0.05)
    stopifnot(anova(lm(m_error~factor(true_alpha), d))[1, "Pr(>F)"] > 0.05)
    stopifnot(anova(lm(m_error~factor(true_I), d))[1, "Pr(>F)"] > 0.05)

    stopifnot(anova(lm(N_error~factor(true_alpha), d))[1, "Pr(>F)"] > 0.05)
    stopifnot(anova(lm(N_error~factor(true_I), d))[1, "Pr(>F)"] > 0.05)
    stopifnot(anova(lm(N_error~true_m, d))[1, "Pr(>F)"] > 0.05)

    alpha.m.cor <- cor.test(d[,alpha_error], d[,true_m], method="spearman")
@

\Cref{fig:abcptm2} shows {\color{red}\cout{\gls{MAP}}}
{\color{blue}\uline{posterior mean}} point estimates of the BA model parameters
obtained with kernel-assisted ABC on simulated data. The estimates shown
correspond only to the simulations where the $m$ parameter was set to 2,
however the results for $m = 3$ and $m = 4$ were similar
(\cref{fig:abcptm3,fig:abcptm4}). Average boundaries of 95\% HPD intervals are
given in \cref{tab:abchpd}.

The accuracy of the parameter estimates obtained with kernel-assisted ABC
paralleled the results from the \gls{kSVR} classifier. Of the four parameters,
$\alpha$ was the most accurately estimated, with point estimates having a
median [IQR] absolute error of 
    \Sexpr{d[,round(median(alpha_error), 2)]} 
    [\Sexpr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(alpha_error, 0.75), 2)]}].
The errors when the true value of $\alpha$ was zero were significantly greater
than those for the other values 
    (Wilcoxon rank-sum test, $p$ = $\Sexpr{latexSN(round(zero.alpha.test$p.value, 4))}$).
Errors in estimating $\alpha$ 
%also varied with the true value of $m$ just at
%the threshold of statistical significance
%    (one-way ANOVA, $p 
%    \Sexpr{pp(anova(lm(alpha_error~true_m, d))[1, "Pr(>F)"], eq=TRUE)}$),
{\color{blue}\uline{did not vary across the true values of $m$ or $I$ (both one-way ANOVA).}} 
Estimates for $I$ were relatively accurate, with point estimate errors of
    \Sexpr{d[,round(median(I_error))]} 
    [\Sexpr{d[,round(quantile(I_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(I_error, 0.75))]}] individuals.
These errors were significantly higher when the true value of $\alpha$ was
at least 1
    (Wilcoxon rank-sum test, $p$ = $\Sexpr{latexSN(round(alpha.I.test$p.value, 4))}$)
and when the true value of $I$ was 2000 ($p < 10^{-5}$). The true value of $m$
did not affect the estimates of $I$ (one-way ANOVA).

The $m$ parameter was estimated correctly in only
    \Sexpr{as.integer(m.tbl[1] * 100)} \%
of simulations {\color{red}\sout{barely better than random guessing}}. The true
values of the other parameters did not significantly affect the estimates of
$m$ (both one-way ANOVA). Finally, the total number of nodes $N$ was
consistently over-estimated by about a factor of two
    (error \Sexpr{d[,round(median(N_error))]} 
    [\Sexpr{d[,round(quantile(N_error, 0.25))]} - 
     \Sexpr{d[,round(quantile(N_error, 0.75))]}] individuals).
No parameters influenced the accuracy of the $N$ estimates (all one-way ANOVA).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{abc-point-estimate-m2}
  \vspace{6pt}
  \caption[
    Posterior mean point estimates for \gls{BA} model parameters obtained by
    running \software{netabc} on simulated data, for simulations with \gls{m} =
    2.
  ]{
    Posterior mean point estimates for \gls{BA} model parameters obtained by
    running \software{netabc} on simulated data, for simulations with \gls{m} =
    2. Dashed lines indicate true values. (A) Estimates of \gls{alpha} and
    \gls{I} which were varied in these simulations against known values. (B)
    Estimates of \gls{m} and \gls{N} which were held fixed in these simulations
    at the values \gls{m} = 2 and \gls{N} = 5000.
  }
  \label{fig:abcptm2}
\end{figure}

@@

%    \Sexpr{d[,median(alpha_error)]} 
%    [\Sexpr{d[,quantile(alpha_error, 0.25)]}-\Sexpr{d[,quantile(alpha_error, 0.75)]}].

@@

%    ($p$ values $\Sexpr{pp(alpha.glm[Parameter == "alpha", min(p)], eq=FALSE)}$
%     $\Sexpr{pp(alpha.glm[Parameter == "m", min(p)], eq=FALSE)}$,

@@

%    (Spearman's $\rho$ = \Sexpr{round(alpha.test$estimate, 2)},
%     $p \Sexpr{pp(alpha.test$p.value, eq=TRUE)}$).

>>>>

@@

To test the full kernel-assisted \gls{ABC} algorithm, we simulated three trees each under
a variety of parameter values, and ran the \software{netabc} program to
estimate posterior distributions for the parameters. The parameter values and
priors used are listed in \cref{tab:abcexpt}. The tree kernel meta-parameters
were set to $\lambda = 0.3$ and $\sigma = 4$. The \gls{SMC} algorithm was run
with 1000 particles, five sampled datasets per particle, and the $\alpha$
parameter (not to be confused with the \gls{BA} preferential attachment
parameter, see \cref{subsec:adaptsmc}) set to 0.95. The algorithm was
stopped when the acceptance rate of the \gls{MH} kernel dropped below 1.5\%,
the same criterion used by \citeauthor{del2012adaptive}. Approximate marginal
posterior densities for each parameter were calculated using the
\software{density} function in \software{R} applied to the final weighted
population of particles. {\color{red}\sout{Credible intervals were obtained for
each parameter using the \software{HPDinterval} function in the \software{coda}
package~\autocite{plummer2006coda}.}} {\color{blue}\uline{Posterior means and
credible intervals were obtained for each parameter using the
\software{wtd.mean} and \software{wtd.quantile} functions in the
\software{Hmisc} package~\autocite{harrell2016hmisc}. }}

@@

Trees simulated under different values of \gls{alpha} were visibly quite
distinct (\cref{fig:alphatrees}). In particular, higher values of \gls{alpha}
produce networks with a small number of highly connected nodes which, once
infected, are likely to transmit to many other nodes. This results in a more
unbalanced, ladder-like structure in the phylogeny, compared to networks with
lower \gls{alpha} values. None of the other three parameters produced trees
which were as easily distinguished from each other
(\cref{fig:Itrees,fig:mtrees,fig:Ntrees,fig:Itrees}).  Sackin's index, which
measures tree imbalance, was significantly correlated with all four parameters
    (for $\alpha$, $I$, $m$, and $N$ respectively: Spearman's rho =
     0.85,
     \ensuremath{-0.12},
     \ensuremath{-0.13},
     0.09;
     $p$-values
     ${<}10^{-5}$,
     $0.003$,
     ${<}10^{-5}$,
     ${<}10^{-5}$).
The ratio of internal to terminal branch lengths was negatively correlated with
\gls{alpha} and \gls{I}, and positively correlated with \gls{m} and \gls{N}
  (Spearman's rho
    \ensuremath{-0.84},
    \ensuremath{-0.69},
    0.1,
    0.18;
  all $p < 10^{-5}$).

@@

@@

When classifying $\alpha$, the \gls{kSVR} classifier had an average $R^2$ of 
    0.92,
compared to 
    0.56
for the \gls{nltt}-based SVR, and
    0.75
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify the $m$ parameter in any epidemic scenario, with average
$R^2$ values of 
  0.12 for \gls{kSVR},
  0.01 for the \gls{nltt}, and
  0.06
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVR} was slightly higher
on 1000-tip trees 
    (average $R^2$ 
     0.01,
     0.11,
     0.32
     for 100, 500, and 1000 tips respectively).

The accuracy of classifiers $I$ varied significantly with the number of tips in
the tree. For 100-tip trees, the average $R^2$ values were
  0.7,
  0.55, and
  0.02
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  0.93,
  0.83, and
  0.07.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  0.08
for the smallest epidemic and smallest sample size, to
  0.82
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  0.01
to
  0.54.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  0.03
and little variation between scenarios.

@@

The accuracy of grid search estimates largely paralleled that of the \gls{kSVR}
classifiers. \Cref{fig:gridest} shows point estimates and 95\% highest density
intervals for each of the \gls{BA} parameters, for one replicate experiment
with 500-tip trees. Plots showing the point estimates for all replicates can be
found in \cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all
parameters except $m$, the error of point estimates was negatively correlated
with the number of sampled tips in the tree (for
\gls{alpha}, \gls{I}, and \gls{N} respectively: Spearman's $\rho$ = 
    \ensuremath{-0.22},
    \ensuremath{-0.51},
    \ensuremath{-0.16};
$p$-values
    $4\!\times\!10^{-4}$,
    ${<}10^{-5}$,
    $0.01$).
The highest density intervals obtained for all parameters were extremely wide,
occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

The \gls{alpha} parameter was the most accurately estimated, with point
estimates having an average deviation of 
    0.14
from the true value, on a grid from 0 to 2. The error of point estimates varied
significantly between true values of \gls{alpha}
    (one-way \gls{ANOVA}, $p$ ${<}10^{-5}$). In
particular, errors were lower for the values \gls{alpha} = 1.0 and 1.25 than
for the other values
    (average errors 
    0.03
    for \gls{alpha} = 1.0 or 1.5 vs.
    0.17
    for \gls{alpha} $\neq$ 1.0 or 1.5),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
     \cref{fig:gridptalpha}).
These two values exhibited different qualitative behaviour than the other
values in terms of the distribution of kernel scores along the grid
(\cref{fig:gridalpha}). In particular, there was a pronounced peak in scores
around the true value, in contrast to the other values where the scores were
flat around the true value. The effect was most obvious for the value
\gls{alpha} = 1.25.

The average absolute error of the point estimates for \gls{I} was 
    310 individuals,
on a grid of 500 to 5000, and these errors differed between true values of
\gls{I}
    (one-way \gls{ANOVA}, $p =0.001$).
The errors for $2000 \leq I \leq 3000$ were higher than those for the other
values
    (average errors
     430
     for $2000 \leq I \leq 3000$ vs.
     250
     for $I < 2000$ or $I > 3000$),
and this difference was significant
    (Wilcoxon rank-sum test, $p =6\!\times\!10^{-4}$,
     \cref{fig:gridptI}).
Kernel score distributions for all test values exhibited a similar rounded
shape (\cref{fig:gridI}).

The average error for \gls{m} was
    1.31 edges per vertex,
on a grid from 1 to 6. The error varied significantly between the true values
of \gls{m} 
    (one-way \gls{ANOVA}, $p {<}10^{-5}$).
Errors for the value \gls{m} = 1 were lower than the other values
    (average errors
    0.1
    for $m = 1$ vs.
    1.55
    for $m > 1$),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
     \cref{fig:gridptm}).
The value $m = 1$ causes the network to take on a distinct shape relative to
higher \gls{m} values, namely a tree (\ie there are no cycles, see
\cref{subsec:treeshape}). The kernel score distribution had a peak at $m = 1$
when this was the true value, and a valley at $m = 1$ when the true value of
$m$ was greater that 1 (\cref{fig:gridm}).

The average error for \gls{N} was 
    2419 individuals,
on a grid from 1000 to 15000, and was varied significantly with the true value
of \gls{N}
    (one-way \gls{ANOVA}, $p {<}10^{-5}$).
The errors were lower for $N \leq 3000$
    (average errors
     740
     for $N \leq 3000$ vs.
     2979
     for $N > 3000$),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
     \cref{fig:gridptN}).
The kernel score distribution had a peak at $N = 1000$ when this was the true
value, and a valley there otherwise (\cref{fig:gridN}). Except for this valley,
the distributions were flat for $N > 3000$.

@@

\Cref{fig:abcptm2} shows \gls{MAP} point estimates of the BA model parameters
obtained with kernel-assisted ABC on simulated data. The estimates shown correspond only
to the simulations where the $m$ parameter was set to 2, however the results
for $m = 3$ and $m = 4$ were similar (\cref{fig:abcptm3,fig:abcptm4}). Average
boundaries of 95\% HPD intervals are given in \cref{tab:abchpd}.

The accuracy of the parameter estimates obtained with kernel-assisted ABC
paralleled the results from the \gls{kSVR} classifier. Of the four parameters,
$\alpha$ was the most accurately estimated, with point estimates having a
median [IQR] absolute error of 
    0.11 
    [0.03 - 
    0.25].
The errors when the true value of $\alpha$ was zero were significantly greater
than those for the other values 
    (Wilcoxon rank-sum test, $p$ = $0$).
Errors in estimating $\alpha$ 
%also varied with the true value of $m$ just at
%the threshold of statistical significance
%    (one-way ANOVA, $p 
%    pp(anova(lm(alpha_error~true_m, d))[1, "Pr(>F)"], eq=TRUE)$),
{\color{blue}\uline{did not vary across the true values of $m$ or $I$ (both one-way ANOVA).}} 
Estimates for $I$ were relatively accurate, with point estimate errors of
    492 
    [294 - 
    782] individuals.
These errors were significantly higher when the true value of $\alpha$ was
at least 1
    (Wilcoxon rank-sum test, $p$ = $0$)
and when the true value of $I$ was 2000 ($p < 10^{-5}$). The true value of $m$
did not affect the estimates of $I$ (one-way ANOVA).

The $m$ parameter was estimated correctly in only
    37 \%
of simulations, barely better than random guessing. The true values of the
other parameters did not significantly affect the estimates of $m$ (both
one-way ANOVA). Finally, the total number of nodes $N$ was consistently
over-estimated by about a factor of two
    (error 4153 
    [3660 - 
     4489] individuals).
No parameters influenced the accuracy of the $N$ estimates (all one-way ANOVA).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{abc-point-estimate-m2}
  \vspace{6pt}
  \caption[
    \Acrlong{MAP} point estimates for \gls{BA} model parameters obtained by
    running \software{netabc} on simulated data, for simulations with \gls{m} =
    2.
  ]{
    \Acrlong{MAP} point estimates for \gls{BA} model parameters obtained by
    running \software{netabc} on simulated data, for simulations with \gls{m} =
    2. Dashed lines indicate true values. (A) Estimates of \gls{alpha} and
    \gls{I} which were varied in these simulations against known values. (B)
    Estimates of \gls{m} and \gls{N} which were held fixed in these simulations
    at the values \gls{m} = 2 and \gls{N} = 5000.
  }
  \label{fig:abcptm2}
\end{figure}

@@

%    d[,median(alpha_error)] 
%    [d[,quantile(alpha_error, 0.25)]-d[,quantile(alpha_error, 0.75)]].

@@

%    ($p$ values $pp(alpha.glm[Parameter == "alpha", min(p)], eq=FALSE)$
%     $pp(alpha.glm[Parameter == "m", min(p)], eq=FALSE)$,

@@

%    (Spearman's $\rho$ = round(alpha.test$estimate, 2),
%     $p pp(alpha.test$p.value, eq=TRUE)$).
